Analytic reproducibility in articles receiving open data badges at Psychological Science: An observational assessment

### 2020-07-07 manuscript/manuscript.Rmd
added references and adjusted some terminology

```{r load_packages}
# load packages
library(knitr) # for literate programming
library(papaja) # for article template
library(tidyverse) # for data munging
library(tidylog) # inline analysis feedback
library(here) # for finding files
library(DescTools) # for calculating confidence intervals
library(english) # for converting numbers to words
#library(metafor) # for meta-analysis
library(meta) # for forest plot
library(osfr) # to download data from the Open Science Framework (for meta-analysis)
```

@@

We employed a retrospective non-comparative case-study design based on @hardwicke2018a. The numerical discrepancy between ‘old’ values reported in the target articles and ‘new’ values obtained in our analyses was quantified using percentage error[^percentError] ($PE = \frac{|new-old|}{old}\times100$). Numerical discrepancies were classified as ‘minor’ (0% > PE < 10%) or ‘major’ (PE >= 10%). If an old *p* value fell on the opposite side of the $\alpha = .05$ boundary relative to a new *p* value we additionally recorded a ‘decision error’. We recorded if there was insufficient information to complete aspects of the analyses.

@@

```{r scatterPlot, fig.cap="Scatterplot showing reanalysis p-values as a function of reported p-values, classified by reproducibility outcome. Axes are on a log scale. Dashed lines represent the typical 0.05 threshold for statistical significance. For display purposes, 41 values below 0.001 are not shown.", fig.path="figs/"}
source(here('analysis','plot_reproScatter.R')) # build plot
scatterPlot # show plot
```
```{r}
source(here('analysis','meta_analysis.R'))
```

@@

After author involvement, `r majorErrors` major numerical discrepancies remained amongst the `r valuesChecked` target values examined across all articles (`r paste0(majorErrors_percent,'% ',majorErrors_percentCI)`). This included `r decisionErrors` decision error for which we obtained a statically significant p value of `r data_values %>% filter(comparisonOutcome =='DECISION_ERROR') %>% pull(obtainedValue)` in contrast to the reported non-significant p-value of `r data_values %>% filter(comparisonOutcome =='DECISION_ERROR') %>% pull(reportedValue)`. A scatterplot illustrating the consistency of old and new p-values is displayed in Figure \@ref(fig:scatterPlot). The frequency of numerical discrepancies by value types is displayed in Figure \@ref(fig:valueTypePlot).

@@

Some reproducibility issues were resolved through involvement of original authors, however, relying on such assistance is not ideal for several reasons: (1) it substantially increases the workload and time investment, both for researchers attempting to re-use data and for original authors; (2) important details about and materials related to the study are more likely to be forgotten or misplaced over time, reducing the ability of authors to assist [@vines2014]; (3) authors may be unwilling or unable to respond, as was the case for three of the articles we examined. Author involvement highlighted that non-reproducibility was often caused by unclear, incomplete, or incorrect reporting of analytic procedures in the original papers, issues which could often be resolved through the provision of additional information or clarifications beyond what was previously reported. In other cases, authors informed us of errors in shared data files or analysis scripts, suggesting that at some stage the original data, analyses, and research report had become decoupled. In some cases, neither we nor the original authors could reproduce the target values and the causal locus of non-reproducibility remained unidentified; it was no longer possible to reconstruct the original analysis pipeline.

>>>>

```{r load_packages}
# load packages
library(knitr) # for literate programming
library(papaja) # for article template
library(tidyverse) # for data munging
library(tidylog) # inline analysis feedback
library(here) # for finding files
library(DescTools) # for calculating confidence intervals
library(english) # for converting numbers to words
library(meta) # for meta-analysis
library(osfr) # to download data from the Open Science Framework (for meta-analysis)
```

@@

We employed a retrospective non-comparative case-study design based on @hardwicke2018a. The numerical discrepancy between **original values** reported in the target articles and **reanalysis values** obtained in our reproducibility checks was quantified using percentage error[^percentError] ($PE = \frac{|reanalysis-original|}{original}\times100$). Numerical discrepancies were classified as ‘minor’ (0% > PE < 10%) or ‘major’ (PE >= 10%). If an old *p* value fell on the opposite side of the $\alpha = .05$ boundary relative to a new *p* value we additionally recorded a ‘decision error’. We recorded if there was insufficient information to complete aspects of the analyses.

@@

```{r scatterPlot, fig.cap="Scatterplot showing reanalysis p-values as a function of original p-values, classified by reproducibility outcome. Axes are on a log scale. Dashed lines represent the typical 0.05 threshold for statistical significance. For display purposes, 41 values below 0.001 are not shown.", fig.path="figs/"}
source(here('analysis','plot_reproScatter.R')) # build plot
scatterPlot # show plot
```
```{r}
source(here('analysis','meta_analysis.R'))
```

@@

After author involvement, `r majorErrors` major numerical discrepancies remained amongst the `r valuesChecked` target values examined across all articles (`r paste0(majorErrors_percent,'% ',majorErrors_percentCI)`). This included `r decisionErrors` decision error for which we obtained a statically significant p value of `r data_values %>% filter(comparisonOutcome =='DECISION_ERROR') %>% pull(obtainedValue)` in contrast to the reported non-significant p-value of `r data_values %>% filter(comparisonOutcome =='DECISION_ERROR') %>% pull(reportedValue)`. A scatterplot illustrating the consistency of original and reanalysis p-values is displayed in Figure \@ref(fig:scatterPlot). The frequency of numerical discrepancies by value types is displayed in Figure \@ref(fig:valueTypePlot).

@@

Some reproducibility issues were resolved through involvement of original authors, however, relying on such assistance is not ideal for several reasons: (1) it substantially increases the workload and time investment, both for researchers attempting to re-use data and for original authors; (2) important details about and materials related to the study are more likely to be forgotten or misplaced over time, reducing the ability of authors to assist [@vines2014; @minocher2020]; (3) authors may be unwilling or unable to respond, as was the case for three of the articles we examined. Author involvement highlighted that non-reproducibility was often caused by unclear, incomplete, or incorrect reporting of analytic procedures in the original papers, issues which could often be resolved through the provision of additional information or clarifications beyond what was previously reported. In other cases, authors informed us of errors in shared data files or analysis scripts, suggesting that at some stage the original data, analyses, and research report had become decoupled. In some cases, neither we nor the original authors could reproduce the target values and the causal locus of non-reproducibility remained unidentified; it was no longer possible to reconstruct the original analysis pipeline.
